---
title: "Rworksheet_Aurelio#3a"
author: "Jeron J. Aurelio"
date: "2025-10-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. There is a built-in vector LETTERS contains the uppercase letters of the alphabet
### and letters which contains the lowercase letters of the alphabet.

```{r} 
alphabet <- LETTERS
alphabet

alphabet_small <- letters
alphabet_small
```
### a. You need to produce a vector that contains the first 11 letters.

```{r}
alphapet <- LETTERS
alphapet[1.11]
```
### b. Produce a vector that contains the odd numbered letters.

```{r}
alphabet <- LETTERS
odd <- alphabet[seq(1,length(alphabet),2)]
odd
```
### c. Produce a vector that contains the vowels

```{r}
alphabet <- LETTERS 
vowels <- alphabet[c(1, 5, 9, 15, 21)]
vowels
```
### d. Produce a vector that contains the last 5 lowercase letters.

```{r}
alphabet <- letters 
last_fivelower <- tail (alphabet, 5)
last_fivelower
```
### e. Produce a vector that contains letters between 15 to 24 letters in lowercase.

``` {r}
lower_alphabet <- letters
lower_alphabet [15:24]
```
### 2. Create a vector(not a dataframe) with the average temperatures in April for Tugue-garao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City. The average

``` {r}
cities <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
temp <- c(42, 39, 34, 34, 30, 27)
cities
temp
```
### a. What is the R code and its result for creating a character vector for the city/town
##of Tuguegarao City, Manila, Iloilo City, Tacloban,Samal Island, and Davao City? Name the object as city. The names should follow the same order as in the instruction.

``` {r}
cities <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
```
### b. The average temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.Name the object as temp. Write the R code and its output. Numbers should also follow what is in the instruction.

``` {r}
temperature <- c(42, 39, 34, 34, 30, 27)
```
### c. Create a dataframe to combine the city and the temp by using ‘data.frame(). What the R code and its result?

```{r}
cities <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
temp <- c(42, 39, 34, 34, 30, 27)
frame = data.frame(cities,temp)
frame
```
### d. Associate the dataframe you have created in 2.(c) by naming the columns usingthe names() function. Change the column names by using names() function as City and Temperature. What is the R code and its result?

```{r}
cities <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
temp <- c(42, 39, 34, 34, 30, 27)
frame = data.frame(cities,temp)
names (frame) <- c("City", "Temperature")
frame
```
### e. Print the structure by using str() function. Describe the output.
```{r}
cities <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
temp <- c(42, 39, 34, 34, 30, 27)
frame = data.frame(cities,temp)
names (frame) <- c("City", "Temperature")
frame

str (frame)
```
### f. From the answer in d, what is the content of row 3 and row 4 What is its R code and its output?

```{r}
cities <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City")
temp <- c(42, 39, 34, 34, 30, 27)
frame = data.frame(cities,temp)
names (frame) <- c("City", "Temperature")
frame[3:4, ]
```
### g. From the answer in d, display the city with highest temperature and the city with the lowest temperature. What is its R code and its output?

```{r}
maximum <- max(frame$Temperature)
minimum <- min(frame$Temperature)

frame[frame$Temperature == maximum,]
frame[frame$Temperature == minimum,]
```
### 2. Create a matrix of one to eight and eleven to fourteen with four columns and three rows.
### a. What will be the R code for the #2 question and its result?
```{r}
matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 11, 12, 13, 14), nrow = 3)
```
### b. Multiply the matrix by two. What is its R code and its result?
```{r}
matrix <- matrix(c(1,2,3,4,5,7,8,9,10,11,12,13,14), nrow =3)
double_matrix <- matrix * 2
double_matrix
```
### c. What is the content of row 2? What is its R code?
```{r}
matrix [2, ]
```
### d. What will be the R code if you want to display the column 3 and column 4 in row 1 and row 2? What is its output?
```{r}
matrix[1:2, 3:4]
```
### e. What is the R code is you want to display only the columns in 2 and 3, row 3? What is its output?

``` {r}
matrix[3, 2:3]
```
### f. What is the R code is you want to display only the columns 4? What is its output?
``` {r}
matrix[ , 4]
```
### g. Name the rows as isa, dalawa, tatlo and columns as uno, dos, tres, quatro
### for the matrix that was created in b.‘. What is its R code and corresponding output?

``` {r}
matrix <- matrix(c(1,2,3,4,5,6,7,8,11,12,13,14),nrow = 3, byrow = TRUE)
rownames(matrix) <- c("isa", "dalawa", "tatlo")
colnames(matrix) <- c("uno", "dos", "tres", "quatro")
matrix
```
### h. From the original matrix you have created in a, reshape the matrix by assigning a
### new dimension with dim(). New dimensions should have 2 columns and 6 rows. What will
### be the R code and its output?

``` {r}
dim(matrix) <- c(6, 2)
matrix
```
### 3. An array contains 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1

``` {r}
array_dta <- c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1)
array_dta
```
### a. Create an array for the above numeric values. Each values will be repeated twice
### What will be the R code if you are to create a three-dimensional array with 4 columns and 2 rows. What will be its output?


```{r} 
num = array(array_dta, dim = c(2,4,3))
num
```

